---
title: "data_challenge2"
author: "Maya Lu, Emaneula Beale,James Lee"
date: "3/6/2021"
output: html_document
---

## 1. SET UP
```{r}
rm(list = ls())
library(tidyr)
library(ggplot2)
library(quanteda)
emails <- read.csv("Clinton.csv")
table(emails$Recipient)

# create normalized version of dtm
dtm <- emails[, 10:3009]
row_lengths = apply(dtm, 1, function(z) sqrt(sum(z^2)))
dtm_norm <- dtm/row_lengths
dtm_norm <- na.omit(dtm_norm)

neg <- unlist(read.table("negative-words.txt", stringsAsFactors = F))
pos <- unlist(read.table("positive-words.txt", stringsAsFactors = F))
```

## 2. DATA PROCESSING
```{r}
# 1) get vector of all words in the DTM
words <- colnames(emails)

# ----------------------------------

# 2) find index's of irrelevant words

# beginning with X (e.g. X2004)
x <- words[grepl("^[xX]", words)]
xIdx <- which(words %in% x) 

# begining with "sent"
sent <- words[grepl("^sent", words)]
sentIdx <- which(words %in% sent)

# stopwords
stopIdx <- which(words %in% stopwords(language = "en"))

# from
from <- words[grepl("^from", words)]
fromIdx <- which(words %in% from) 

# subject
subject <- words[grepl("subject", words)]
subjectIdx <- which(words %in% subject) 

# cc
cc <- words[grepl("cc_", words)]
ccIdx <- which(words %in% cc) 

# state
state <- words[grepl("^state_", words)]
stateIdx <- which(words %in% state) 

# release
release <- words[grepl("^release_in", words)]
releaseIdx <- which(words %in% release) 

# message
message <- words[grepl("message", words)]
messageIdx <- which(words %in% message) 

# full
full <- words[grepl("full_", words)]
fullIdx <- which(words %in% full) 

# department
department <- words[grepl("department", words)]
departmentIdx <- which(words %in% department) 

# -----------------------------------

# 3) Subset email
emails_clean <- subset(emails, select = -c(xIdx, sentIdx, stopIdx, 
                                           fromIdx, subjectIdx, ccIdx, stateIdx, 
                                           releaseIdx, messageIdx, fullIdx,
                                           departmentIdx))

# ---------------------------------
# 4) Create year indicator column for emails_clean
emails_clean <-separate(emails_clean, DateSent, c("year", "month") )
dtm_clean <- emails_clean[-c(1:10)]

#save(emails_clean, dtm_clean, file = "processedDTM.Rdata")

```


## Building functions

### 1) define function to display  frequent words
```{r}

clusterFreq <- function(k, dat) {
  out <- kmeans(dat, centers = k, nstart = 3)
  
  # create freq words dataframe
  freqWords <- matrix(NA, 10, k)
  freqWords <- data.frame(freqWords)
  
  for (i in 1:k) {
    clu <- i
    centroid.i <- out$centers[clu,]
    freqWords[,i] <- names(sort(centroid.i, decreasing = T)[1:10])
  }
  # return data frame
  return(freqWords)
}
```

### 2) Function that displays distinct words
```{r}

clusterDistinct <- function(k, dat)  {
  out <- kmeans(dat, centers = k, nstart = 3)
  
  # create distinct words dataframe
  distinctWords <- matrix(NA, 10, k)
  distinctWords <- data.frame(distinctWords)
  
  # populate data frame
  for (i in 1:k) {
    clu <- i
    centroid.i <- out$centers[clu,]
    centroid.noti <- apply(out$centers[-clu,], 2, mean)
    distinctWords[,i] <- names(sort(centroid.i - centroid.noti, decreasing = T)[1:10])
  }
  
  # return data frame
  return(distinctWords)
}
```

### 3) function that outputs sentiment ratio 
```{r}
calcSentiment <- function(dat) {
  countPos = rep(NA, nrow(dat))
  countNeg = rep(NA, nrow(dat))
  
  for (i in 1:nrow(dat)) {
    wordsIdx = which(t(dat[i,]) > 1)
    wordNames <- colnames(dat)
    words <- wordNames[wordsIdx] 
    
    countPos[i] = sum(as.numeric(words %in% pos))
    countNeg[i] = sum(as.numeric(words %in% neg))
  }
  return(sum(countPos)/sum(countPos, countNeg))
}
```


### 4) Sentiment Regression Analysis function
```{r}
calcSentimentRegression <- function(keyword, dat) {
  countPos = rep(NA, nrow(dat))
  countNeg = rep(NA, nrow(dat))
  keyword.counts = dat[,keyword]
  
  for (i in 1:nrow(dat)) {
    wordsIdx = which(t(dat[i,]) > 1)
    wordNames <- colnames(dat)
    words <- wordNames[wordsIdx] 
    
    countPos[i] = sum(as.numeric(words %in% pos))
    countNeg[i] = sum(as.numeric(words %in% neg))
  }
  
  df <- data.frame(keyword = keyword.counts, countPos = countPos, countNeg = countNeg)
  df$ratio <- df$countPos/(df$countPos + df$countNeg)
  df$ratio[df$countNeg == 0 & df$countPos == 0] <- 0.5
  
  return(lm(ratio ~ keyword, df))
}
```

### 5) function that gets number of documents containing asia
```{r}
# =====================
# CALC NDOCS PER REGION
# =====================
# function that gets number of documents containing asia
# @param: dat - dataframe to be analyzed
# @param: region - vector of keywords for each region

nDocs <- function(dat, region) {
  idx <- which(colnames(dat) %in% region)
  df <- dat[ ,idx]
  df$total <- rowSums(df)
  
  # num documents containing regional keywords
  df$has_region <- df$total > 0
  ndocs <- sum(df$has_region)
  
  return(ndocs)
}

```


### 6) function that calculates sentiment of a region
```{r}
# =========================
# CALC SENTIMENT FOR REGION
# =========================
# function that calculates sentiment of a region
# @param: dataframe
# @param: region - vector of keywords for region
# @returns: ratio of positive to positive & negative words for dataframe

regionSentiment <- function(dat, region) {
  # create regional df
  idx <- which(colnames(dat) %in% region)
  df <- dat[ ,idx]
  df$total <- rowSums(df)
  df$has_region <- df$total > 0
  
  # get index of observations w/ dataframe
  idx <- which(df$has_region == 1)
  df <- dat[idx,]
  sent <- calcSentiment(df) #helper function to calculate sentiment of df
  
  return(sent)
}

# similar function for any keyword.
keywordSentiment <- function(dat, keyword) {
  # create regional df
  idx <- which(colnames(dat) %in% keyword)
  df <- dat[ ,idx]
  df$total <- rowSums(df)
  df$has_keyword <- df$total > 0
  
  # get index of observations w/ dataframe
  idx <- which(df$has_keyword == 1)
  df <- dat[idx,]
  sent <- calcSentiment(df) #helper function to calculate sentiment of df
  
  return(sent)
}
```


## Data Analysis

### K-CLUSTERS OVER TIME
```{r}
# create dataframes for each respective years - UNIGRAMS
dtm_2009 <- emails_clean[emails_clean$year == "2009", ][11:2786]
dtm_2010 <- emails_clean[emails_clean$year == "2010", ][11:2786]
dtm_2011 <- emails_clean[emails_clean$year == "2011", ][11:2786]
dtm_2012 <- emails_clean[emails_clean$year == "2012", ][11:2786]

# import distinct words function from functions.R
# find distinct words from 2009 to 2012
dw2009 <- clusterDistinct(10, dtm_2009)
fw2009 <- clusterFreq(10, dtm_2009)

dw2010 <- clusterDistinct(10, dtm_2010)
fw2009 <- clusterFreq(10, dtm_2009)

dw2011 <- clusterDistinct(10, dtm_2011)
fw2011 <- clusterFreq(10, dtm_2011)

dw2012 <- clusterDistinct(10, dtm_2012)
fw2012 <- clusterDistinct(10, dtm_2012)

save(dw2009, dw2010, dw2011, dw2012, file = "topicsOverTime.Rdata")
```


### SENTIMENT OVER TIME
```{r}
overall2009sentiment = calcSentiment(dtm_2009)
overall2010sentiment = calcSentiment(dtm_2010)
overall2011sentiment = calcSentiment(dtm_2011)
overall2012sentiment = calcSentiment(dtm_2012)
overallSents <- c(overall2009sentiment,overall2010sentiment,overall2011sentiment,overall2012sentiment )

```


### Sentiment over time - by Recipients
```{r}
abedin <- c("huma", "abedin")
mills <- c("cheryl", "mills")
sullivan <- c("jacob", "sullivan")
jiloty <- c("lauren", "jiloty")

# NUMBER DOCUMENTS 
# Huma Abedin
nAbedin <- nDocs(dtm_clean, abedin)

# Cheryl Mills
nMills <- nDocs(dtm_clean, mills) 

# Jacob Sullivan
nSullivan <- nDocs(dtm_clean, sullivan)

# Lauren Jiloty
nJiloty <- nDocs(dtm_clean, jiloty) 

nRecipDocs <- c(nAbedin, nMills, nSullivan, nJiloty)
barplot(nRecipDocs, 
        names.arg = c("Abdein", "Mills", "Sullivan", "Jiloty"),
        main = "Number of Documents Mentioning Each recipient")

# ----------------------------------
# VISUALIZE PROPORTION OF DOCUMENTS FOR EACH REGION ACROSS TIME
par(mfrow = c(2, 2))

# abedin
abedinDocs2009 <- nDocs(dtm_2009, abedin)/nrow(dtm_2009)
abedinDocs2010 <- nDocs(dtm_2010, abedin)/nrow(dtm_2010)
abedinDocs2011 <- nDocs(dtm_2011, abedin)/nrow(dtm_2011)
abedinDocs2012 <- nDocs(dtm_2012, abedin)/nrow(dtm_2012)
abedinDocs <- c(abedinDocs2009, abedinDocs2010, abedinDocs2011, abedinDocs2012)
barplot(abedinDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Huma Abedin")

# mills
millsDocs2009 <- nDocs(dtm_2009, mills)/nrow(dtm_2009)
millsDocs2010 <- nDocs(dtm_2010, mills)/nrow(dtm_2010)
millsDocs2011 <- nDocs(dtm_2011, mills)/nrow(dtm_2011)
millsDocs2012 <- nDocs(dtm_2012, mills)/nrow(dtm_2012)
millsDocs <- c(millsDocs2009, millsDocs2010, millsDocs2011, millsDocs2012)
barplot(millsDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Cheryl Mills")

# sullivan
sullivanDocs2009 <- nDocs(dtm_2009, sullivan)/nrow(dtm_2009)
sullivanDocs2010 <- nDocs(dtm_2010, sullivan)/nrow(dtm_2010)
sullivanDocs2011 <- nDocs(dtm_2011, sullivan)/nrow(dtm_2011)
sullivanDocs2012 <- nDocs(dtm_2012, sullivan)/nrow(dtm_2012)
sullivanDocs <- c(sullivanDocs2009, sullivanDocs2010, sullivanDocs2011, sullivanDocs2012)
barplot(sullivanDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Jacob Sullivan")

# jiloty
jilotyDocs2009 <- nDocs(dtm_2009, jiloty)/nrow(dtm_2009)
jilotyDocs2010 <- nDocs(dtm_2010, jiloty)/nrow(dtm_2010)
jilotyDocs2011 <- nDocs(dtm_2011, jiloty)/nrow(dtm_2011)
jilotyDocs2012 <- nDocs(dtm_2012, jiloty)/nrow(dtm_2012)
jilotyDocs <- c(jilotyDocs2009, jilotyDocs2010, jilotyDocs2011, jilotyDocs2012)
barplot(jilotyDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Lauren Jiloty")

# --------------------------

# Abedin
AbedinSent2009 <- keywordSentiment(dtm_2009, abedin)
AbedinSent2010 <- keywordSentiment(dtm_2010, abedin)
AbedinSent2011 <- keywordSentiment(dtm_2011, abedin)
AbedinSent2012 <- keywordSentiment(dtm_2012, abedin)
AbedinSents <- c(AbedinSent2009, AbedinSent2010, AbedinSent2011, AbedinSent2012)

# Mills
millsSent2009 <- keywordSentiment(dtm_2009, mills)
millsSent2010 <- keywordSentiment(dtm_2010, mills)
millsSent2011 <- keywordSentiment(dtm_2011, mills)
millsSent2012 <- keywordSentiment(dtm_2012, mills)
millsSents <- c(millsSent2009, millsSent2010, millsSent2011, millsSent2012)

# sullivan
sullivanSent2009 <- keywordSentiment(dtm_2009, sullivan)
sullivanSent2010 <- keywordSentiment(dtm_2010, sullivan)
sullivanSent2011 <- keywordSentiment(dtm_2011, sullivan)
sullivanSent2012 <- keywordSentiment(dtm_2012, sullivan)
sullivanSents <- c(sullivanSent2009, sullivanSent2010, sullivanSent2011, sullivanSent2012)

# jiloty
jilotySent2009 <- keywordSentiment(dtm_2009, jiloty)
jilotySent2010 <- keywordSentiment(dtm_2010, jiloty)
jilotySent2011 <- keywordSentiment(dtm_2011, jiloty)
jilotySent2012 <- keywordSentiment(dtm_2012, jiloty) # this turns out to be NA so set it to 0.5
jilotySent2012 <- 0.5
jilotySents <- c(jilotySent2009, jilotySent2010, jilotySent2011, jilotySent2012)

# --------------------

# VISUALIZE SENTIMENT OVER TIME
par(mfrow = c(1,1))
#library(ggplot2)
recipientSents <- data.frame(AbedinSents, millsSents, sullivanSents, jilotySents)

ggplot(dat = recipientSents, aes(x = c(1,2,3,4), y = AbedinSents, col = "Huma Abedin")) + 
  geom_line() +
  geom_line(aes(y = millsSents, col = "Cheryl Mills")) +
  geom_line(aes(y = sullivanSents, col = "Jacob Sullivan")) +
  geom_line(aes(y = jilotySents, col = "Lauren Jiloty")) +
  geom_line(aes(y = overallSents, col = "overall"), linetype = "dashed", col = "black") +
  xlab("year (2009-2012)") +
  ylab("Positive Sentiment Ratio") +
  ggtitle("Sentiment Towards Recipients Across Years")+
  theme(plot.title = element_text(size=12, hjust = 0.5, face = "bold"), 
        panel.background = element_rect(fill = "lightgrey", colour = "lightgrey",
                                        size = 2, linetype = "solid"))
```


### Sentiment over time - by region
```{r}
# REGIONAL ANALYSIS OVER TIME

asia <- c("china", "southeast", "asia", "korea", "japan", "singapore", "india", "indonesia")
middle_east <- c("syria", "gaza", "qatar", "turkey", "egypt", "jordan", "pakistan", "palestine", "saudi", "yemen", "israel", "iran", "iraq")
africa <- c("africa" , "kenya", "congo", "morocco", "uganda", "libya", "guinea", "sudan")
la_carib <- c("latin","caribbean", "brazil", "colombia", "venezuela", "cuba", "mexico", "haiti", "honduras")
europe <- c("germany", "greece", "ireland", "russia","uk","armenia",  "poland", "spain", "italy")

# ---------------
# NUMBER DOCUMENTS 
# ASIA
nAsiaDoc <- nDocs(dtm_clean, asia)

# MIDDLE EAST
nMeDoc <- nDocs(dtm_clean, middle_east)

# AFRICA
nAfricaDoc <- nDocs(dtm_clean, africa) # 592
 
# LATIN AMERICA & CARIBBEAN
nLaCaribDoc <- nDocs(dtm_clean, la_carib) # 854

# EUROPE
nEuropeDoc <- nDocs(dtm_clean, europe) # 690

# -----------------------------
# VISUALIZE FREQUENCY
nRegionDocs <- c(nAsiaDoc, nMeDoc, nAfricaDoc, nLaCaribDoc, nEuropeDoc)
barplot(nRegionDocs, 
        names.arg = c("Asia", "Middle East", "Africa", "Latin America/Carib", "Europe"),
        main = "Number of Documents Mentioning Each Region")

# ----------------------------------
# VISUALIZE PROPORTION OF DOCUMENTS FOR EACH REGION ACROSS TIME
par(mfrow = c(3, 2))

# ASIA
asiaDocs2009 <- nDocs(dtm_2009, asia)/nrow(dtm_2009)
asiaDocs2010 <- nDocs(dtm_2010, asia)/nrow(dtm_2010)
asiaDocs2011 <- nDocs(dtm_2011, asia)/nrow(dtm_2011)
asiaDocs2012 <- nDocs(dtm_2012, asia)/nrow(dtm_2012)
asiaDocs <- c(asiaDocs2009, asiaDocs2010, asiaDocs2011, asiaDocs2012)
barplot(asiaDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Asia")

# MIDDLE EAST
meDocs2009 <- nDocs(dtm_2009, middle_east)/nrow(dtm_2009)
meDocs2010 <- nDocs(dtm_2010, middle_east)/nrow(dtm_2010)
meDocs2011 <- nDocs(dtm_2011, middle_east)/nrow(dtm_2011)
meDocs2012 <- nDocs(dtm_2012, middle_east)/nrow(dtm_2012)
meDocs <- c(meDocs2009, meDocs2010, meDocs2011, meDocs2012)
barplot(meDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Middle East")

# AFRICA
africaDocs2009 <- nDocs(dtm_2009, africa)/nrow(dtm_2009)
africaDocs2010 <- nDocs(dtm_2010, africa)/nrow(dtm_2010)
africaDocs2011 <- nDocs(dtm_2011, africa)/nrow(dtm_2011)
africaDocs2012 <- nDocs(dtm_2012, africa)/nrow(dtm_2012)
africaDocs <- c(africaDocs2009, africaDocs2010, africaDocs2011, africaDocs2012)
barplot(africaDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Africa")

# LATIN AMERICA & CARIBBEAN
laCaribDocs2009 <- nDocs(dtm_2009, la_carib)/nrow(dtm_2009)
laCaribDocs2010 <- nDocs(dtm_2010, la_carib)/nrow(dtm_2010)
laCaribDocs2011 <- nDocs(dtm_2011, la_carib)/nrow(dtm_2011)
laCaribDocs2012 <- nDocs(dtm_2012, la_carib)/nrow(dtm_2012)
laCaribDocs <-c(laCaribDocs2009, laCaribDocs2010, laCaribDocs2011, laCaribDocs2012)
barplot(laCaribDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Latin America/Carib")

# EUROPE
europeDocs2009 <- nDocs(dtm_2009, europe)/nrow(dtm_2009)
europeDocs2010 <- nDocs(dtm_2010, europe)/nrow(dtm_2010)
europeDocs2011 <- nDocs(dtm_2011, europe)/nrow(dtm_2011)
europeDocs2012 <- nDocs(dtm_2012, europe)/nrow(dtm_2012)
europeDocs <- c(europeDocs2009, europeDocs2010, europeDocs2011, europeDocs2012)
barplot(europeDocs,
        names.arg = c("2009", "2010", "2011", "2012"), 
        main = "Proportion of Documents Pertaining to Europe")

# --------------------------

# REGIONAL SENTIMENT OVER TIME

# ASIA
asiaSent2009 <- regionSentiment(dtm_2009, asia)
asiaSent2010 <- regionSentiment(dtm_2010, asia)
asiaSent2011 <- regionSentiment(dtm_2011, asia)
asiaSent2012 <- regionSentiment(dtm_2012, asia)
asiaSents <- c(asiaSent2009, asiaSent2010, asiaSent2011, asiaSent2012)

# MIDDLE EAST
meSent2009 <- regionSentiment(dtm_2009, middle_east)
meSent2010 <- regionSentiment(dtm_2010, middle_east)
meSent2011 <- regionSentiment(dtm_2011, middle_east)
meSent2012 <- regionSentiment(dtm_2012, middle_east)
meSents <- c(meSent2009, meSent2010, meSent2011, meSent2012 )

# AFRICA
africaSent2009 <- regionSentiment(dtm_2009, africa)
africaSent2010 <- regionSentiment(dtm_2010, africa)
africaSent2011 <- regionSentiment(dtm_2011, africa)
africaSent2012 <- regionSentiment(dtm_2012, africa)
africaSents <- c(africaSent2009, africaSent2010, africaSent2011, africaSent2012)

# LATIN AMERICA & CARIBBEAN
laCaribSent2009 <- regionSentiment(dtm_2009, la_carib)
laCaribSent2010 <- regionSentiment(dtm_2010, la_carib)
laCaribSent2011 <- regionSentiment(dtm_2011, la_carib)
laCaribSent2011 <- 0.5
laCaribSent2012 <- regionSentiment(dtm_2012, la_carib)
laCaribSents <- c(laCaribSent2009, laCaribSent2010, laCaribSent2011, laCaribSent2012 )

# EUROPE
europeSent2009 <- regionSentiment(dtm_2009, europe)
europeSent2010 <- regionSentiment(dtm_2010, europe)
europeSent2011 <- regionSentiment(dtm_2011, europe)
europeSent2012 <- regionSentiment(dtm_2012, europe)
europeSents <- c(europeSent2009, europeSent2010, europeSent2011, europeSent2012)

# --------------------

# VISUALIZE SENTIMENT OVER TIME

par(mfrow = c(1,1))
regionalSents <- data.frame(asiaSents, meSents, africaSents, laCaribSents, europeSents, overallSents)

ggplot(dat = regionalSents, aes( x = c(1,2,3,4), y = asiaSents, col = "asia")) + 
  geom_line() +
  geom_line(aes(y = meSents, col = "middle east")) + 
  geom_line(aes(y = africaSents, col = "africa")) +
  geom_line(aes(y = laCaribSents, col = "latin america/caribbean")) +
  geom_line(aes(y = europeSents, col = "europe")) +
  geom_line(aes(y = overallSents, col = "overall"), linetype = "dashed", col = "gray") +
  xlab("year (2009-2012)") +
  ylab("Positive Sentiment Ratio") +
  ggtitle("Sentiment Towards Regions Across Years")+
  ggtitle("Sentiment Towards Recipients Across Years")+
  theme(plot.title = element_text(size=12, hjust = 0.5, face = "bold"), 
        panel.background = element_rect(fill = "lightgrey", colour = "lightgrey",
                                        size = 2, linetype = "solid"))

```

